---
title: "Association Analysis - Final Project"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(regclass)
#setwd("your path")

##Data Cleaning Code From Raw Data##
#country_data = read.csv("country_data.csv")
#new_country_data <- country_data[, c("name","homicide_rate","gdp",  "urban_population_growth", #"pop_growth", "internet_users", "gdp_per_capita", "refugees")]
#clean_data_homcides <- new_country_data[!is.na(new_country_data$homicide_rate), ]
#no_dupes <- clean_data_homcides%>% distinct(name, .keep_all = TRUE)
#final_data <- no_dupes
final_data <- read.csv("Final Data.csv")

##Knit Document To View The Analysis##

```

***Visual analysis using scatterplots:***

**#1. Homicide Rate vs GDP**
```{r, warning=FALSE}
ggplot(final_data, aes(y = homicide_rate, x = gdp))+
  geom_point()
```

**Perform log transformation on x and y due to heavy skewness**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = log(homicide_rate + 1), x = log(gdp + 1)))+
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

**Plot seems to indicate an approximate moderate negative linear correlation. Use Pearson's correlation for testing.**


**2. Homicide Rate vs Urban Population Growth**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = homicide_rate, x=urban_population_growth))+
  geom_point()
```

**Perform log transformation on y due to nonlinear and non-monotonic relationship**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = log(homicide_rate + 1), x=urban_population_growth+1))+
  geom_point() +
  geom_smooth(method = "lm", se=FALSE)
```

**Plot can be described with a weak positive linear correlation. Use Pearson's correlation for testing.**


**3. Homicide Rate vs Population Growth**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = homicide_rate, x=pop_growth))+
  geom_point()
```

**Perform log transformation on y due to nonlinear and non-monotonic relationship**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = log(homicide_rate + 1), x=pop_growth))+
  geom_point() +
  geom_smooth(method = "lm", se=FALSE)
```

**Plot can be described with a weak positive linear correlation. Use Pearson's correlation for testing.**


**4. Homicide Rate vs Internet Users Percentage**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = homicide_rate, x=internet_users))+
  geom_point()
```

**Perform log transformation on y due to nonlinearity and non-monotonic relationship**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = log(homicide_rate+1), x=internet_users))+
  geom_point() +
  geom_smooth(method="lm",se=FALSE)
```

**Plot can be described with a moderate negative linear correlation. Use Pearson's correlation for testing.**


**5. Homicide Rate vs GDP Per Capita**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = homicide_rate, x = gdp_per_capita))+
  geom_point()
```

**Perform log transformation on x and y due to heavy skewness**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = log(homicide_rate + 1), x = log(gdp_per_capita + 1)))+
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

**Plot seems to indicate a moderate negative linear correlation. Use Pearson's correlation for testing.**


**6. Homicide Rate vs Refugees**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = homicide_rate, x=refugees))+
  geom_point()
```

**Perform log transformation on x and y due to heavy skewness.**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = log(homicide_rate+1), x=log(refugees)))+
  geom_point() +
  geom_smooth(method="lm",se=FALSE)
```

**Plot seems to indicate a weak negative linear correlation. Use Pearson's correlation for testing.**


***Association Analysis***

**1. log Homicide Rate vs log GDP**
```{r, fig.show='hide', warning=FALSE}
associate(log(homicide_rate + 1)~log(gdp + 1), data=final_data, seed=9750)

```
**There exists a statistically significant correlation, p-value = 0. There is a moderate negative linear correlation, r = -0.2951072**


**2. log Homicide Rate vs Urban Population Growth**
```{r, fig.show='hide', warning=FALSE}
associate(log(homicide_rate+1)~urban_population_growth, data=final_data, seed=9750)
```
**There does not exist a statistically significant correlation, p-value = 0.186**


**3. log Homicide Rate vs Population Growth**
```{r, fig.show='hide', warning=FALSE}
associate(log(homicide_rate+1)~pop_growth, data=final_data, seed=9750)
```
**There is not a statistically significant correlation, p-value = 0.268**


**4. log Homicide Rate vs Internet Users Percentage**
```{r, fig.show='hide', warning=FALSE}
associate(log(homicide_rate+1)~internet_users, data=final_data, seed=9750)
```
**There exists a statistically significant correlation, p-value = 0. There is a moderate negative linear correlation, r = -0.3205548**


**5. log Homicide Rate vs log GDP Per Capita**
```{r, fig.show='hide', warning=FALSE}
associate(log(homicide_rate+1)~log(gdp_per_capita+1), data=final_data, seed=9750)
```
**There exists a statistically significant association, p-value = 0. There is a moderate negative linear correlation, r = -0.3025433**


**6. log Homicide Rate vs log Refugee Population**
```{r, fig.show='hide', warning=FALSE}
associate(log(homicide_rate+1)~log(refugees+1), data=final_data, seed=9750)
```
**There is not a statistically significant correlation, p-value = 0.456**
