---
title: "Full Code"
output: word_document
date: "2024-12-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(regclass)
setwd("C:/Users/Taylor's Home/OneDrive - The City University of New York/Fall 2024/STA 9750 Tools/STA 9750 Final Project/Global Data")
country_data = read.csv("country_data.csv")
```

```{r cleaning Data}
new_country_data = country_data[, c("name","homicide_rate","gdp",  "urban_population_growth", "pop_growth", "internet_users", "gdp_per_capita", "refugees")]
head(new_country_data)

clean_data_homcides <- new_country_data[!is.na(new_country_data$homicide_rate), ]
no_dupes = clean_data_homcides%>% distinct(name, .keep_all = TRUE)

final_data = no_dupes
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
#library(dplyr)
#library(ggplot2)
#library(regclass)
#setwd("your path")
#country_data = read.csv("country_data.csv")
new_country_data <- country_data[, c("name","homicide_rate","gdp",  "urban_population_growth", "pop_growth", "internet_users", "gdp_per_capita", "refugees")]
clean_data_homcides <- new_country_data[!is.na(new_country_data$homicide_rate), ]
no_dupes <- clean_data_homcides%>% distinct(name, .keep_all = TRUE)
final_data <- no_dupes
```

***Visual analysis using scatterplots:***

**#1. Homicide Rate vs GDP**
```{r, warning=FALSE}
ggplot(final_data, aes(y = homicide_rate, x = gdp))+
  geom_point()
```

**Perform log transformation on x and y due to heavy skewness**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = log(homicide_rate + 1), x = log(gdp + 1)))+
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

**Plot seems to indicate an approximate moderate negative linear correlation. Use Pearson's correlation for testing.**


**2. Homicide Rate vs Urban Population Growth**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = homicide_rate, x=urban_population_growth))+
  geom_point()
```

**Perform log transformation on y due to nonlinear and non-monotonic relationship**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = log(homicide_rate + 1), x=urban_population_growth+1))+
  geom_point() +
  geom_smooth(method = "lm", se=FALSE)
```

**Plot can be described with a weak positive linear correlation. Use Pearson's correlation for testing.**


**3. Homicide Rate vs Population Growth**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = homicide_rate, x=pop_growth))+
  geom_point()
```

**Perform log transformation on y due to nonlinear and non-monotonic relationship**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = log(homicide_rate + 1), x=pop_growth))+
  geom_point() +
  geom_smooth(method = "lm", se=FALSE)
```

**Plot can be described with a weak positive linear correlation. Use Pearson's correlation for testing.**


**4. Homicide Rate vs Internet Users Percentage**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = homicide_rate, x=internet_users))+
  geom_point()
```

**Perform log transformation on y due to nonlinearity and non-monotonic relationship**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = log(homicide_rate+1), x=internet_users))+
  geom_point() +
  geom_smooth(method="lm",se=FALSE)
```

**Plot can be described with a moderate negative linear correlation. Use Pearson's correlation for testing.**


**5. Homicide Rate vs GDP Per Capita**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = homicide_rate, x = gdp_per_capita))+
  geom_point()
```

**Perform log transformation on x and y due to heavy skewness**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = log(homicide_rate + 1), x = log(gdp_per_capita + 1)))+
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

**Plot seems to indicate a moderate negative linear correlation. Use Pearson's correlation for testing.**


**6. Homicide Rate vs Refugees**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = homicide_rate, x=refugees))+
  geom_point()
```

**Perform log transformation on x and y due to heavy skewness.**
```{r, warning=FALSE, message=FALSE}
ggplot(final_data, aes(y = log(homicide_rate+1), x=log(refugees)))+
  geom_point() +
  geom_smooth(method="lm",se=FALSE)
```

**Plot seems to indicate a weak negative linear correlation. Use Pearson's correlation for testing.**


***Association Analysis***

**1. log Homicide Rate vs log GDP**
```{r, fig.show='hide', warning=FALSE}
associate(log(homicide_rate + 1)~log(gdp + 1), data=final_data, seed=9750)

```
**There exists a statistically significant correlation, p-value = 0. There is a moderate negative linear correlation, r = -0.2951072**


**2. log Homicide Rate vs Urban Population Growth**
```{r, fig.show='hide', warning=FALSE}
associate(log(homicide_rate+1)~urban_population_growth, data=final_data, seed=9750)
```
**There does not exist a statistically significant correlation, p-value = 0.186**


**3. log Homicide Rate vs Population Growth**
```{r, fig.show='hide', warning=FALSE}
associate(log(homicide_rate+1)~pop_growth, data=final_data, seed=9750)
```
**There is not a statistically significant correlation, p-value = 0.268**


**4. log Homicide Rate vs Internet Users Percentage**
```{r, fig.show='hide', warning=FALSE}
associate(log(homicide_rate+1)~internet_users, data=final_data, seed=9750)
```
**There exists a statistically significant correlation, p-value = 0. There is a moderate negative linear correlation, r = -0.3205548**


**5. log Homicide Rate vs log GDP Per Capita**
```{r, fig.show='hide', warning=FALSE}
associate(log(homicide_rate+1)~log(gdp_per_capita+1), data=final_data, seed=9750)
```
**There exists a statistically significant association, p-value = 0. There is a moderate negative linear correlation, r = -0.3025433**


**6. log Homicide Rate vs log Refugee Population**
```{r, fig.show='hide', warning=FALSE}
associate(log(homicide_rate+1)~log(refugees+1), data=final_data, seed=9750)
```
**There is not a statistically significant correlation, p-value = 0.456**
```{r ANOVA}
# Perform a multiple ANOVA test
anova_model = aov(homicide_rate ~ urban_population_growth + pop_growth + internet_users + gdp + refugees + gdp_per_capita, data = final_data)

# Summarize the ANOVA results
summary(anova_model)

```

```{r}
library(car)
# Fit a linear model
linear_model = lm(homicide_rate ~ gdp + urban_population_growth + pop_growth + internet_users + gdp_per_capita + refugees, data = final_data)

# Calculate VIF
vif_values = vif(linear_model)
print(vif_values)

cor_matrix = cor(final_data[, c("gdp", "urban_population_growth", "pop_growth", "internet_users", "gdp_per_capita", "refugees")])
print(cor_matrix)

# Scale numeric variables to standardize them
numeric_data = final_data[, c("gdp", "urban_population_growth", "pop_growth", "internet_users", "gdp_per_capita", "refugees")]
scaled_data = scale(numeric_data)

# Calculate condition number
condition_number = kappa(cor(scaled_data, use = "complete.obs"))
print(condition_number)

```

```{r Regressional Analysis}
library(ggplot2)
#country_data = read.csv("country_data.csv")
new_country_data = country_data[, c("name","homicide_rate","gdp",  "urban_population_growth", "pop_growth", "internet_users", "gdp_per_capita", "refugees")]
head(new_country_data)
clean_data_homcides <- new_country_data[!is.na(new_country_data$homicide_rate), ]
no_dupes = clean_data_homcides%>% distinct(name, .keep_all = TRUE)
final_data = no_dupes
names(final_data)
#Regression Analysis
##Fit a Regression line(HOMICIDE RATE vs GDP)
M<-lm(homicide_rate~gdp,data=final_data)
##Scatterplot with line(HOMICIDE RATE vs GDP)
library(ggplot2)
ggplot(data=final_data, aes(x = gdp, y = homicide_rate)) +
  geom_point() +
  scale_x_log10() +
  #geom_smooth(method = "lm", se = FALSE, col = "red") +
  labs(x = "GDP", y = "Homicide Rate") 

##Coefficients,standard error,R^2,RMSE,p-value(HOMICIDE RATE vs GDP)
summary(M)

##sum of squared error reduction(HOMICIDE RATE vs GDP)
anova(M)

##Confidence interval for coefficients(HOMICIDE RATE vs GDP)
confint(M,level=0.95)

##Fit a Regression line(HOMICIDE RATE vs urban_population_growth)
M<-lm(homicide_rate~urban_population_growth,data=final_data)
##Scatterplot with line(HOMICIDE RATE vs urban_population_growth)
library(ggplot2)
ggplot(data=final_data, aes(x = urban_population_growth, y = homicide_rate)) +
  geom_point() +
  #scale_x_log10() +
  #geom_smooth(method = "lm", se = FALSE, col = "red") +
  labs(x = "urban_population_growth", y = "homicide_rate") 

##Coefficients,standard error,R^2,RMSE,p-value(HOMICIDE RATE vs urban_population_growth)
summary(M)

##sum of squared error reduction(HOMICIDE RATE vs urban_population_growth)
anova(M)

##Confidence interval for coefficients(HOMICIDE RATE vs urban_population_growth)
confint(M,level=0.95)

##Fit a Regression line(HOMICIDE RATE vs pop_growth)
M<-lm(homicide_rate~pop_growth,data=final_data)
##Scatterplot with line(HOMICIDE RATE vs pop_growth)
library(ggplot2)
ggplot(data=final_data, aes(x = pop_growth, y = homicide_rate)) +
  geom_point() +
  #scale_x_log10() +
  #geom_smooth(method = "lm", se = FALSE, col = "red") +
  labs(x = "pop_growth", y = "homicide_rate") 

##Coefficients,standard error,R^2,RMSE,p-value(HOMICIDE RATE vs pop_growth)
summary(M)

##sum of squared error reduction(HOMICIDE RATE vs pop_growth)
anova(M)

##Confidence interval for coefficients(HOMICIDE RATE vs pop_growth)
confint(M,level=0.95)

##Fit a Regression line(HOMICIDE RATE vs internet_users)
M<-lm(homicide_rate~internet_users,data=final_data)
##Scatterplot with line(HOMICIDE RATE vs internet_users)
library(ggplot2)
ggplot(data=final_data, aes(x = internet_users, y = homicide_rate)) +
  geom_point() +
  #scale_x_log10() +
  #geom_smooth(method = "lm", se = FALSE, col = "red") +
  labs(x = "internet_users", y = "homicide_rate") 

##Coefficients,standard error,R^2,RMSE,p-value(HOMICIDE RATE vs internet_users)
summary(M)

##sum of squared error reduction(HOMICIDE RATE vs internet_users)
anova(M)

##Confidence interval for coefficients(HOMICIDE RATE vs internet_users)
confint(M,level=0.95)

##Fit a Regression line(HOMICIDE RATE vs gdp_per_capita)
M<-lm(homicide_rate~gdp_per_capita,data=final_data)
##Scatterplot with line(HOMICIDE RATE vs gdp_per_capita)
library(ggplot2)
ggplot(data=final_data, aes(x = gdp_per_capita, y = homicide_rate)) +
  geom_point() +
  scale_x_log10() +
  #geom_smooth(method = "lm", se = FALSE, col = "red") +
  labs(x = "log(gdp_per_capita)", y = "homicide_rate") 

##Coefficients,standard error,R^2,RMSE,p-value(HOMICIDE RATE vs gdp_per_capita)
summary(M)

##sum of squared error reduction(HOMICIDE RATE vs gdp_per_capita)
anova(M)

##Confidence interval for coefficients(HOMICIDE RATE vs gdp_per_capita)
confint(M,level=0.95)

##Fit a Regression line(HOMICIDE RATE vs refugees)
M<-lm(homicide_rate~refugees,data=final_data)
##Scatterplot with line(HOMICIDE RATE vs refugees)
library(ggplot2)
ggplot(data=final_data, aes(x = refugees, y = homicide_rate)) +
  geom_point() +
  scale_x_log10() +
  #geom_smooth(method = "lm", se = FALSE, col = "red") +
  labs(x = "refugees", y = "homicide_rate") 

##Coefficients,standard error,R^2,RMSE,p-value(HOMICIDE RATE vs refugees)
summary(M)

##sum of squared error reduction(HOMICIDE RATE vs refugees)
anova(M)

##Confidence interval for coefficients(HOMICIDE RATE vs refugees)
confint(M,level=0.95)

#Fit a Multilinear Regression Model
M<-lm(homicide_rate~gdp+urban_population_growth+pop_growth+internet_users+gdp_per_capita+refugees,data=final_data)

##Scatterplot with line(HOMICIDE RATE vs all predictors)
library(ggplot2)
ggplot(data=final_data, aes(x = gdp+urban_population_growth+pop_growth+internet_users+gdp_per_capita+refugees, y = homicide_rate)) +
  geom_point() +
  scale_x_log10() +
  geom_smooth(method = "lm", se = FALSE, col = "red") +
  labs(x = "All predictors", y = "homicide_rate") 

##Coefficients,standard error,R^2,RMSE,p-value(HOMICIDE RATE vs all predictors)
summary(M)

##sum of squared error reduction(HOMICIDE RATE vs all predictors)
anova(M)

##Confidence interval for coefficients(HOMICIDE RATE vs all predictors)
confint(M,level=0.95)

```